---
title: "The difference in mass between subspecies of common chaffinch."
author: "Emma Rand"
engine: knitr
# format:
#   html
format:
  pdf:
    documentclass: article
    papersize: A4
    mainfont: Calibri
    geometry:
      - top=20mm
      - left=25mm
      - right=25mm
execute:
  echo: false
  include: true
  error: false
  message: false
  warning: false
bibliography: references.bib
---

```{r}
options(digits = 2)
```

# Introduction

A number of subspecies of the [common
chaffinch](https://en.wikipedia.org/wiki/Common_chaffinch), *Fringilla
coelebs*, have been described based principally on the differences in
the pattern and colour of the adult male plumage [@suárez2009]. Two of
groups of these subspecies are the "coelebs group" (@fig-coelebs) that
occurs in Europe and Asia and the "canariensis group" (@fig-canariensis) that occurs on the Canary Islands.

::: {#fig-subspecies layout-ncol="2"}
![*F. c.
coelebs*](images/coelebs.jpg){#fig-coelebs}

![*F. c.
palmae*](images/canariensis.jpg){#fig-canariensis}

**Adult male *Fringilla coelebs***. Coelebs group on the left (Andreas
Trepte, CC BY-SA 2.5 <https://creativecommons.org/licenses/by-sa/2.5>,
via Wikimedia Commons) and canariensis group on the right (H.
Zell, CC BY-SA 3.0 <https://creativecommons.org/licenses/by-sa/3.0>, via
Wikimedia Commons).
:::

# Methods

```{r}
#| label: load-packages
library(tidyverse)
```

```{r}
#| label: import-data
file <- "data-raw/chaff.txt"
chaff <- read_table(file)
```

```{r}
#| label: data-summary
chaff_summary <- chaff |> 
  group_by(subspecies) |>  
  summarise(mean = mean(mass, na.rm = TRUE),
            sd = sd(mass, na.rm = TRUE),
            n = length(mass),
            se = sd / sqrt(n))
```

```{r}
#| label: extract-summary
# extract values for inline reporting

# sample sizes
n_coelebs <- chaff_summary |> 
  filter(subspecies == "coelebs") |> 
  pull(n)

n_canariensis <- chaff_summary |> 
  filter(subspecies == "canariensis") |> 
  pull(n)

```

We randomly sampled `r n_coelebs` *F. c. coelebs* males and
`r n_canariensis` *F. c. palmae* males and determined their mass with
spring scales. Analysis was carried out with R [@r-core] and tidyverse
packages [@tidyverse] and the manuscript was prepared with Quarto
[@Allaire_Quarto_2024].

# Results

Canariensis have higher mass than coelebs. See @fig-chaff.

::: {#fig-chaff}
```{r}
#| fig-width: 3
#| fig-height: 3
#| fig-align: left
ggplot() +
  geom_point(data = chaff, aes(x = subspecies, y = mass),
             position = position_jitter(width = 0.1, height = 0),
             colour = "gray50") +
  geom_errorbar(data = chaff_summary, 
                aes(x = subspecies, ymin = mean - se, ymax = mean + se),
                width = 0.3) +
  geom_errorbar(data = chaff_summary, 
                aes(x = subspecies, ymin = mean, ymax = mean),
                width = 0.2) +
  scale_y_continuous(name = "Mass (g)", 
                     limits = c(0, 30), 
                     expand = c(0, 0)) +
  scale_x_discrete(name = "Subspecies", 
                   labels = c("Canariensis", "Coelebs")) +
  theme_classic()
```

**Canariensis chaffinches are heavier than Coelebs chaffinches**. The
mean mass of `r n_coelebs` randomly sampled males from each subspecies
was determined. Error bars means are $\pm$ 1 standard error. Data 
analysis was conducted in R [@r-core] with tidyverse [@tidyverse] 
packages .
:::

The distribution mass can be can been seen in @fig-histograms for
Canariensis in @fig-hist-canariensis and for Coelebs in
@fig-hist-coelebs. We don't really need more figures for this data
but it allows you to see how to create multipanel figures.

::: {#fig-histograms layout-ncol="2"}
```{r}
#| label: fig-hist-canariensis
chaff |> 
  filter(subspecies == "canariensis") |>
  ggplot(aes(x = mass)) +
  geom_histogram(bins = 8) +
  scale_x_continuous(name = "Mass (g)", 
                     expand = c(0, 0)) +
  scale_y_continuous(name = "Count", 
                     expand = c(0, 0)) +
  theme_classic()

```

```{r}
#| label: fig-hist-coelebs
chaff |> 
  filter(subspecies == "coelebs") |>
  ggplot(aes(x = mass)) +
  geom_histogram(bins = 8) +
  scale_x_continuous(name = "Mass (g)", 
                     expand = c(0, 0)) +
  scale_y_continuous(name = "Count", 
                     expand = c(0, 0)) +
  theme_classic()

```


**Distribution of mass for two subspecies of chaffinch.** @fig-hist-canariensis, *F. c. canariensis* 
 and, @fig-hist-coelebs, *F. c. coelebs*.  The mean mass of `r n_coelebs` randomly sampled males from each subspecies was determined.
:::

A summary of the results are gratuitously reproduced in @tbl-chaff just 
to include a table.

```{r}
#| label: tbl-chaff
chaff_summary |> 
  knitr::kable(caption = "Summary of mass(g) for the two subspecies of chaffinch.",
               col.names = names(chaff_summary) |> str_to_sentence())

```

# Discussion

yadda yadda

# References
