---
title: "The difference in mass between subspecies of common chaffinch."
author: "Emma Rand"
engine: knitr
format:
  html
execute:
  echo: false
  include: true
  error: false
  message: false
  warning: false
---





# Introduction

A number of subspecies of the [common chaffinch]( https://en.wikipedia.org/wiki/Common_chaffinch), 
*Fringilla coelebs*, have been described based principally on 
the differences in the pattern and colour of the adult male plumage. 
Two of groups of these subspecies are the "coelebs group" that occurs 
in Europe and Asia and the "canariensis group" that occurs on
the Canary Islands. 



# Methods
```{r}
#| label: load-packages
library(tidyverse)
```

```{r}
#| label: import-data
file <- "data-raw/chaff.txt"
chaff <- read_table(file)
```

```{r}
#| label: data-summary
chaff_summary <- chaff |> 
  group_by(subspecies) |>  
  summarise(mean = mean(mass, na.rm = TRUE),
            sd = sd(mass, na.rm = TRUE),
            n = length(mass),
            se = sd / sqrt(n))
```


```{r}
#| label: extract-summary
# extract values for inline reporting

# sample sizes
n_coelebs <- chaff_summary |> 
  filter(subspecies == "coelebs") |> 
  pull(n)

n_canariensis <- chaff_summary |> 
  filter(subspecies == "canariensis") |> 
  pull(n)

```


We randomly sampled `r n_coelebs` *F. c. coelebs* males and
`r n_canariensis` *F. c. palmae* males and determined their mass with
spring scales.  Analysis was carried out with R and tidyverse packages.


# Results

Canariensis have higher mass than coelebs. See @fig-chaff

::: {#fig-chaff}


```{r}
#| fig-width: 3.5
#| fig-height: 3.5
#| fig-align: left
ggplot() +
  geom_point(data = chaff, aes(x = subspecies, y = mass),
             position = position_jitter(width = 0.1, height = 0),
             colour = "gray50") +
  geom_errorbar(data = chaff_summary, 
                aes(x = subspecies, ymin = mean - se, ymax = mean + se),
                width = 0.3) +
  geom_errorbar(data = chaff_summary, 
                aes(x = subspecies, ymin = mean, ymax = mean),
                width = 0.2) +
  scale_y_continuous(name = "Mass (g)", 
                     limits = c(0, 30), 
                     expand = c(0, 0)) +
  scale_x_discrete(name = "Subspecies", 
                   labels = c("Canariensis", "Coelebs")) +
  theme_classic()
```

**Canariensis chaffinches are heavier than Coelebs chaffinches**. The
mean mass of `r n_coelebs` randomly sampled males from each subspecies was determined. Error bars are $\pm$ 1 standard error

:::


::: {#fig-histograms layout-ncol=2}

```{r}
#| label: fig-hist-canariensis
chaff |> 
  filter(subspecies == "canariensis") |>
  ggplot(aes(x = mass)) +
  geom_histogram(bins = 8) +
  theme_minimal()

```


```{r}
#| label: fig-hist-coelebs
chaff |> 
  filter(subspecies == "coelebs") |>
  ggplot(aes(x = mass)) +
  geom_histogram(bins = 8) +
  theme_minimal()

```

**Distribution of mass for *F. c. canariensis* and *F. c. coelebs***. The
mean mass of `r n_coelebs` randomly sampled males from each subspecies was determined.


:::

# Discussion


# References

,