---
title: "The difference in mass between subspecies of common chaffinch."
author: "Emma Rand"
engine: knitr
format:
  html
execute:
  echo: false
  include: true
  error: false
  message: false
  warning: false
bibliography: references.bib
---

```{r}
options(digits = 2)
```

# Introduction

A number of subspecies of the [common
chaffinch](https://en.wikipedia.org/wiki/Common_chaffinch), *Fringilla
coelebs*, have been described based principally on the differences in
the pattern and colour of the adult male plumage [@suárez2009]. Two of
groups of these subspecies are the "coelebs group" that occurs in Europe
and Asia and the "canariensis group" that occurs on the Canary Islands.

# Methods

```{r}
#| label: load-packages
library(tidyverse)
```

```{r}
#| label: import-data
file <- "data-raw/chaff.txt"
chaff <- read_table(file)
```

```{r}
#| label: data-summary
chaff_summary <- chaff |> 
  group_by(subspecies) |>  
  summarise(mean = mean(mass, na.rm = TRUE),
            sd = sd(mass, na.rm = TRUE),
            n = length(mass),
            se = sd / sqrt(n))
```

```{r}
#| label: extract-summary
# extract values for inline reporting

# sample sizes
n_coelebs <- chaff_summary |> 
  filter(subspecies == "coelebs") |> 
  pull(n)

n_canariensis <- chaff_summary |> 
  filter(subspecies == "canariensis") |> 
  pull(n)

```

We randomly sampled `r n_coelebs` *F. c. coelebs* males and
`r n_canariensis` *F. c. palmae* males and determined their mass with
spring scales. Analysis was carried out with R [@r-core] and tidyverse
packages [@tidyverse]

# Results

Canariensis have higher mass than coelebs. See @fig-chaff

::: {#fig-chaff}
```{r}
#| fig-width: 3.5
#| fig-height: 3.5
#| fig-align: left
ggplot() +
  geom_point(data = chaff, aes(x = subspecies, y = mass),
             position = position_jitter(width = 0.1, height = 0),
             colour = "gray50") +
  geom_errorbar(data = chaff_summary, 
                aes(x = subspecies, ymin = mean - se, ymax = mean + se),
                width = 0.3) +
  geom_errorbar(data = chaff_summary, 
                aes(x = subspecies, ymin = mean, ymax = mean),
                width = 0.2) +
  scale_y_continuous(name = "Mass (g)", 
                     limits = c(0, 30), 
                     expand = c(0, 0)) +
  scale_x_discrete(name = "Subspecies", 
                   labels = c("Canariensis", "Coelebs")) +
  theme_classic()
```

**Canariensis chaffinches are heavier than Coelebs chaffinches**. The
mean mass of `r n_coelebs` randomly sampled males from each subspecies
was determined. Error bars are $\pm$ 1 standard error
:::

The distribution mass can be can been seen in @fig-histograms for
Canariensis in @fig-hist-canariensis and for Coelebs in
@fig-hist-coelebs.

::: {#fig-histograms layout-ncol="2"}
```{r}
#| label: fig-hist-canariensis
chaff |> 
  filter(subspecies == "canariensis") |>
  ggplot(aes(x = mass)) +
  geom_histogram(bins = 8) +
  theme_minimal()

```

```{r}
#| label: fig-hist-coelebs
chaff |> 
  filter(subspecies == "coelebs") |>
  ggplot(aes(x = mass)) +
  geom_histogram(bins = 8) +
  theme_minimal()

```

![*F. c.
palmae*](images/canariensis.jpg){#fig-canariensis}

![*F. c.
coelebs*](images/coelebs.jpg){#fig-coelebs}



**Distribution of mass for *F. c. canariensis* (@fig-hist-canariensis and @fig-canariensis) and *F. c. coelebs* (@fig-hist-coelebs and @fig-coelebs).**
The mean mass of `r n_coelebs` randomly sampled males from each
subspecies was determined.
:::

These results are gratuitously reproduced in @tbl-chaff just to include
a table.

```{r}
#| label: tbl-chaff
chaff_summary |> 
  knitr::kable(caption = "Summary of mass(g) for the two subspecies of chaffinch.",
               col.names = names(chaff_summary) |> str_to_sentence())

```

# Discussion

# References
